---
title: "HymHub: analysis demo"
output: html_document
---

[HymHub](http://brendelgroup.github.io/HymHub) provides descriptive statistics of several genome features for 13 species in the order Hymenoptera.
These summary statistics are stored in tabular format (tab-separated values or TSV) to facilitate loading into popular statistics packages for analysis.

This notebook isn't designed to do anything particularly useful other than provide a brief demo for loading a data set into an R data frame and pulling data of interest out of that data frame.
Here we will focus on interval loci (iLoci), but everything in this demo also applies to the data tables for other genomic features available from HymHub.

## Read data into memory

In R, tabular data is read into memory using the `read.table` function.
This creates a data frame object, which we can peek at using the `head` function.

```{r}
loci <- read.table("iloci.tsv", header=TRUE, sep="\t", stringsAsFactors=FALSE)
head(loci, n=5) # Let's check out the first 5 rows of the data frame
```

We can use the `str` command for a more detailed description.

```{r}
str(loci)
```

## Get the number of iLoci for each species

iLocus data for all 13 species in HymHub are included in the same table.
The **Species** column is provided to help distinguish which species a particular iLocus belongs to (see [this page](http://brendelgroup.github.io/HymHub/genomes.html) for a full description).
We can grab the species values and determine how many rows (iLoci) correspond to each species.

```{r}
loci.rle <- rle(as.vector(loci$Species))
nloci <- data.frame(loci.rle$values, loci.rle$lengths)
colnames(nloci) <- c("Species", "NumLoci")
nloci
```

# Plot distribution of iLocus GC content

Here, we discard the 10% most extreme iLoci in terms of length, and then plot the GC content of the remaining iLoci by species.

```{r}
# Discard extremely short and extremely long iLoci
lq <- quantile(loci$Length, probs=c(0.05, 0.95))
loci.s <- loci[loci$Length > lq[1] & loci$Length < lq[2],]
lq

# Grab species list
species <- sort(unique(loci$Species))
nspecies <- length(species)

# Plot histograms as lines
h <- hist(loci.s$GCContent[loci.s$Species == species[1]], plot=FALSE, breaks=20)
plot(h$mids, h$counts, type="l", main="", xlab="%GC Content", ylab="Frequency",
     xlim=c(0, 0.8), ylim=c(0, 12000), col=rainbow(nspecies)[1])
for(i in 2:nspecies)
{
  h <- hist(loci.s$GCContent[loci.s$Species == species[i]], plot=FALSE, breaks=20)
  lines(h$mids, h$counts, col=rainbow(nspecies)[i])
}
```
